---
title: "Rmarkdown Lab Notebook Tutorial"
author: "Mark Wadsworth"
date: "6/12/2018"
output: 
    html_document: #This renders an html file. If a PDF is desired use pdf_document
        css: custom.css #This is a css script I wrote to float the table of contents a the top left corner of the document 
        toc: true #Add table of contents
        code_folding: show #Allow code to seen. If hide is used the code is hidden. It can be toggled by clicking the hide/show button
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(pheatmap)
suppressMessages(library(DESeq2)) #The suppressMessages function silences any std out and std err produced by the loading of the package
suppressMessages(library(samr))
suppressMessages(library(digest))
```

## Version Information

This prints out the version of each package that I will use.

```{r collapse=T} 
print(R.version.string)
print( paste("SAMR version:", packageVersion('samr')) )
print( paste("DESeq2 version:", packageVersion('DESeq2')) )
print( paste("pheatmap version:", packageVersion('pheatmap')) )
```

In Rmd you can add in bash or python commands by puting {bash} or {python} at the beginning of your code block.

```{bash, collapse=T}
python --version
bash --version
```

## Load Data

I'm loading TCGA BRCA RNASeq data for TNBC patients.

### read.table

This is the standard way to load data frames in R. I reorder the columns in the data file and put both files into one data structure to easily keep track of my data.

```{r warning=F, message=F}
start = proc.time()
rna = NULL
rna$Data = read.table("TCGA.BRCA.TNBC.log2.normalized.counts.txt", header=T, row.names = 1, sep="\t") #reads in the normalized counts data frame
rna$info = read.table("TCGA.BRCA.TNBC.Phenos.txt", header=T, row.names = 1, sep="\t") #reads in the sample phenotype data frame

rna$Data = rna$Data[,rownames(rna$info)] #This puts the data columns in the same order as the sample info file rows.

print(proc.time() - start)
```

### fread

fread is a faster way to read in files into the same data structure. 

```{r warning=F, message=F}
suppressMessages(library(data.table))
start = proc.time()

rna=NULL
rna$Data=fread("TCGA.BRCA.TNBC.log2.normalized.counts.txt",sep="\t",header=T, data.table = F)
rownames(rna$Data) = rna$Data[,1]
rna$Data = rna$Data[,-1]

rna$info=fread("TCGA.BRCA.TNBC.Phenos.txt",sep="\t",header=T, data.table = F)
rownames(rna$info) = rna$info[,1]
rna$info = rna$info[,-1]

rna$Data = rna$Data[,rownames(rna$info)] #This puts the data columns in the same order as the sample info file rows.

print(proc.time() - start)
```


